---
title: "Reports in Rmarkdown"
author: "Justin and Arturo"
output: pdf_document
---

```{r setup, echo = F, include = F}
library(pander)
require(tidyverse)

#Global Parameters for all code chunks
knitr::opts_chunk$set(cache = TRUE, echo = F)

#When using the control file this param needs to be commented/removed
species_param <- "versicolor"
```


#Main Title

## Report for Species `r species_param`

In this Rmarkdown we will show how to build a workflow around RMarkdown that allows for the automated generation of reports.

##Repository Organisation  

There are several key files that are often useful when writing remports in `RMarkdown` (especially when the project is _big_):  

+ **RMarkdown document** (this one)  
+ **Data** (usually excel/csv files)  
+ **Control file**: used to automatically generate reports  
+ R Scripts and additional data wrangling (helps keep long code out of this document)  
+ Git/version control (we can leave this for another time)

##Rmarkdown Basics  

RMarkdown is a document written in the markdown language that has embedded `R code`. For instance, I can add a calculation into the the text, such as the sum of two integers (three and five): `r 3+5`. We can also add _code chunks_:  

```{r}
table(iris$Species) %>% pander()
```

<!-- We can add latex commands through the text section, such as \newpage -->
\newpage

##Graphical Presentation of Data  

Ultimately, your report will begin to blend text, code, tables and figures. In this case we can include a scatter splot that shows the relationship between sepal length and sepal width for each species:    

```{r, fig.width = 6, fig.height=4}
iris %>% 
  ggplot(aes(x = Sepal.Length, 
                    y = Sepal.Width,
                    fill = Species))+
  geom_point(shape = 21)+
  geom_smooth(aes(colour = Species), method = "lm", se = F)+
  theme_classic()
```
  
\newpage  

##Generating multiple plots  

Now imagine that we want to look at the correlation between sepal lengths and sepal widths for each of the iris plant species, **BUT** we want to keep the other two species unidentifiable... For instance, imagine we want to make a report for _setosa_ but we don't want the person reading it to know what the correlation is for the other two species. 

If you wanted to make three figures for each species, you could make a loop within this RMarkdown: 

```{r, fig.width=6, fig.height=2.25, fig.align='center',results='asis'}
#Loop for unique species names
for (spec in unique(iris$Species)){
 plot<- ggplot(data = iris %>% filter(Species == spec), #only plot filled circles for species of interest
         aes(x = Sepal.Length, 
                    y = Sepal.Width))+
  geom_point(data =  iris, fill = "grey90", shape = 21)+
  geom_point(aes(fill = Species), shape = 21, size = 2)+
  geom_smooth(colour = "tomato", 
              method = "lm", 
              se = F)+
   scale_fill_manual(values = "tomato")+
   ggtitle(spec)+
  theme_classic()
 
 # Page layout
 print(plot)
}
```

\newpage  

##Generating multiple reports  

But, we are interested in generating multiple reports, so the loop should be placed outside of this RMarkdown, for this purpose we make a control file. The following plot will than render for each unique species  

```{r, fig.width = 6, fig.height=4}
ggplot(data = iris %>% filter(Species == species_param), #only plot filled circles for species of interest
         aes(x = Sepal.Length, 
                    y = Sepal.Width))+
  geom_point(data =  iris, fill = "grey90", shape = 21)+
  geom_point(aes(fill = Species), shape = 21, size = 2)+
  geom_smooth(colour = "tomato", 
              method = "lm", 
              se = F)+
   scale_fill_manual(values = "tomato")+
  ggtitle(paste0("Relationship for Species ", species_param))+
  theme_classic()
```
  
  
**Finally**, we can customise large sections of text that someone has written externally and placed in a csv. To do that we map a _csv_ file with the custom text to the specific report parameter (Species). The following is taken from the wikipedia entries for each species:  

```{r echo = F, results='asis'}
Species_description <- read.csv('Species_description.csv', stringsAsFactors = F)
Species_description <- Species_description %>% filter(Species == species_param)
cat(Species_description$Text_string)
```





